## Load Packages & Data
```{r}
library("ggplot2")
library("ggmap")
library("plyr")
library("rgdal")
library("scales")
library("Cairo")
library("maptools")
library("rgeos")
library("dplyr")
library("grid")
library("gridExtra")
library("RColorBrewer")

# TKA data
data <- read.csv("WorkDf5.csv", sep=";")
# State FIP code data
statecodes <- read.delim("state.txt", sep="|", header=T)
# US county map drawing data
counties <- readOGR(dsn = "cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
```

## Clean Data

Create new opioid variables
```{r}
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
data$MEDperDay <- data$MEDperMonth/(365/12)
data$chronic <- ifelse(data$TimeBefore > 2 & data$MEDperDay >= 5, 1, 0)
data$highMED <- ifelse(data$MEDperDay > 100, 1, 0)
data$user <- ifelse(data$MEDperDay > 0,1,0)
```
Remove spurious height and weight values, then create BMI
```{r}
data$ht <- ifelse(data$ht < 48 | data$ht > 86, NA, data$ht)
data$wt <- ifelse(data$wt < 80 | data$wt > 600, NA, data$wt)
data <- mutate(data, BMI = (wt/2.2)/(ht*2.54/100)^2)
```

## Create Table 1
Results for all patients
```{r}
# Age
mean(data$Age)
sd(data$Age)
# BMI
mean(data$BMI, na.rm=T)
sd(data$BMI, na.rm=T)
# Male %
nrow(data[data$gender == "M",])/ nrow(data)
# CKD
sum(data$CKD, na.rm=T)/ nrow(data)
# DM
sum(data$Diabetes, na.rm=T)/ nrow(data)
# HTN
sum(data$HTN, na.rm=T)/ nrow(data)
# OSA
sum(data$OSA, na.rm=T)/ nrow(data)
# PTSD
sum(data$PTSD, na.rm=T)/ nrow(data)
```
Results for non-users
```{r}
# Subset to non-users
data2 <- filter(data, MEDperDay == 0)
# Age
mean(data2$Age)
sd(data2$Age)
# BMI
mean(data2$BMI, na.rm=T)
sd(data2$BMI, na.rm=T)
# Male %
aMale <- nrow(data2[data2$gender == "M",])/ nrow(data2)
aMale
# CKD
aCKD <- sum(data2$CKD, na.rm=T)/ nrow(data2)
aCKD
# DM
aDM <- sum(data2$Diabetes, na.rm=T)/ nrow(data2)
aDM
# HTN
aHTN <- sum(data2$HTN, na.rm=T)/ nrow(data2)
aHTN
# OSA
aOSA <- sum(data2$OSA, na.rm=T)/ nrow(data2)
aOSA
# PTSD
aPTSD <- sum(data2$PTSD, na.rm=T)/ nrow(data2)
aPTSD
```
Results for non-chronic opioid users
```{r}
# Subset
data2 <- filter(data, MEDperDay > 0 & chronic == 0)
# Age
mean(data2$Age)
sd(data2$Age)
# BMI
mean(data2$BMI, na.rm=T)
sd(data2$BMI, na.rm=T)
# Male %
bMale <- nrow(data2[data2$gender == "M",])/ nrow(data2)
bMale
# CKD
bCKD <- sum(data2$CKD, na.rm=T)/ nrow(data2)
bCKD
# DM
bDM <- sum(data2$Diabetes, na.rm=T)/ nrow(data2)
bDM
# HTN
bHTN <- sum(data2$HTN, na.rm=T)/ nrow(data2)
bHTN
# OSA
bOSA <- sum(data2$OSA, na.rm=T)/ nrow(data2)
bOSA
# PTSD
bPTSD <- sum(data2$PTSD, na.rm=T)/ nrow(data2)
bPTSD
```
Results for chronic users
```{r}
# Subset
data2 <- filter(data, chronic == 1)
# Age
mean(data2$Age)
sd(data2$Age)
# BMI
mean(data2$BMI, na.rm=T)
sd(data2$BMI, na.rm=T)
# Male %
cMale <- nrow(data2[data2$gender == "M",])/ nrow(data2)
cMale
# CKD
cCKD <- sum(data2$CKD, na.rm=T)/ nrow(data2)
cCKD
# DM
cDM <- sum(data2$Diabetes, na.rm=T)/ nrow(data2)
cDM
# HTN
cHTN <- sum(data2$HTN, na.rm=T)/ nrow(data2)
cHTN
# OSA
cOSA <- sum(data2$OSA, na.rm=T)/ nrow(data2)
cOSA
# PTSD
cPTSD <- sum(data2$PTSD, na.rm=T)/ nrow(data2)
cPTSD
```
p values
```{r}
# number of non-users
a <- nrow(filter(data, MEDperDay == 0))
# number of non-chronic users
b <- nrow(filter(data, MEDperDay > 0 & chronic == 0))
# number of chronic users
c <- nrow(filter(data, chronic == 1))

# Male %
prop.test(x=c(aMale*a,bMale*b,cMale*c), n=c(a, b, c))
# CKD
prop.test(x=c(aCKD*a,bCKD*b,cCKD*c), n=c(a, b, c))
# DM
prop.test(x=c(aDM*a,bDM*b,cDM*c), n=c(a, b, c))
# HTN
prop.test(x=c(aHTN*a,bHTN*b,cHTN*c), n=c(a, b, c))
# OSA
prop.test(x=c(aOSA*a,bOSA*b,cOSA*c), n=c(a, b, c))
# PTSD
prop.test(x=c(aPTSD*a,bPTSD*b,cPTSD*c), n=c(a, b, c))

# Group patients by use
data$usergrp <- ifelse(data$user == 0, 0, 1)
data$usergrp <- ifelse(data$chronic == 1, 2, data$usergrp)

# Age ANOVA
fit1 <- aov(data$Age ~ data$usergrp)
summary(fit1)

# BMI ANOVA
fit2 <- aov(data$BMI ~ data$usergrp)
summary(fit2)
```

## Create Table 2

Stratify patients by year
```{r}
datagrp <- group_by(data, TKAYear)
data$TKAYear <- as.factor(data$TKAYear)
```
Results for all patients
```{r}
table1 <- summarise(datagrp, count = n(), users = sum(user, na.rm=T)/n(), medMED = median(MEDperDay, na.rm=T), IQR25 = quantile(MEDperDay, na.rm=T)[2], IQR75 = quantile(MEDperDay, na.rm=T)[4], chronic = sum(chronic, na.rm=T)/n(), hundred = sum(highMED, na.rm=T)/n())
table1
```
Results for chronic users
```{r}
datagrp2 <- filter(datagrp, chronic == 1)
table2 <- summarise(datagrp2, chron_med_MED = median(MEDperDay, na.rm=T), IQR25 = quantile(MEDperDay, na.rm=T)[2], IQR75 = quantile(MEDperDay, na.rm=T)[4] )
table2
```
P values
```{r}
# Surgeries/ year
fit1 <- lm(table1$count ~ table1$TKAYear)
summary(fit1)
# Opioid users/ year
fit2 <- lm(table1$users ~ table1$TKAYear)
summary(fit2)
# Median MEDs among all users/ year
fit3 <- lm(table1$medMED ~ table1$TKAYear)
summary(fit3)
# Chronic users/ year
fit4 <- lm(table1$chronic ~ table1$TKAYear)
summary(fit4)
# Median MEDs among chronic users/ year
fit5 <- lm(table2$chron_med_MED ~ table1$TKAYear)
summary(fit5)
# Percent of patients using 100+ MEDs/ year
fit6 <- lm(table1$hundred ~ table1$TKAYear)
summary(fit6)
```

## Bar Graph
```{r}
# Group by division
datagrp3 <- group_by(data, DIVISION)

# For each division, find count, % using opioids, % chronic, % >100 MEDs, and median MEDs 
barData1 <- summarize(datagrp3, N = n(), Users = 100*sum(user, na.rm=T)/n(), Chronic = 100*sum(chronic, na.rm=T)/n(), HighMED = 100*sum(highMED, na.rm=T)/n(), Median = median(MEDperDay, na.rm=T))
# Remove divisions with less than 50 patients
barData1 <- barData1[barData1$N > 49,]

# Subset with just chronic users
datagrp4 <- group_by(data[data$chronic == 1,], DIVISION)
# Find the median MEDs/ day in chronic users
barData2 <- summarize(datagrp4, N = n(), Chron_Median = median(MEDperDay, na.rm=T))
# Remove divisions with < 25 patients
barData2 <- barData2[barData2$N > 24,]

# Create a data frame for graph a
barDF1 <- data.frame(Outcome = c(rep("Opioid Users", 2), rep("Chronic Opioid\nUsers", 2), rep(">100 MEDs\nPer Day", 2)), Group = c(rep(c("5th %", "95th %"),3)))
barDF1$Outcome <- factor(barDF1$Outcome, levels=barDF1$Outcome)
# Add in values for the 5th%ile and 95th%ile divisions for % users, % chronic users, and % >100 MEDs
barDF1$percent <- c(quantile(barData1$Users, c(.05,.95)), quantile(barData1$Chronic, c(.05,.95)), quantile(barData1$HighMED, c(.05,.95)))

# Create a data frame for graph b
barDF2 <- data.frame(Outcome = c(rep("Overall\nMedian MEDs", 2), rep("Chronic User\nMedian MEDs", 2)), Group = c(rep(c("5th %ile Divsion", "95th %ile Division"),2)))
barDF2$Outcome <- factor(barDF2$Outcome, levels=barDF2$Outcome)
# Add in values for the 5th%ile and 95th%ile divisions for median MEDs and chronic user median MEDs
barDF2$median <- c(quantile(barData1$Median, c(.05,.95)), quantile(barData2$Chron_Median, c(.05,.95)))

# Plot a
g1 <- ggplot(barDF1, aes(Outcome, percent, fill=Group)) + geom_bar(position="dodge", stat="identity", color = "black") + theme_bw(base_size=18) + scale_fill_brewer() + labs(x="", y ="Percent") + theme(legend.title=element_blank()) + coord_cartesian(ylim=c(0,70)) + scale_y_continuous(breaks=seq(0,70,10)) + guides(fill=FALSE) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

# Plot b
g2 <- ggplot(barDF2, aes(Outcome, median, fill=Group)) + geom_bar(position="dodge", stat="identity", color = "black") + theme_bw(base_size=18) + coord_cartesian(ylim=c(0,30)) + scale_fill_brewer() + labs(x="", y ="Milligrams") + theme(legend.title=element_blank()) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + guides(fill = guide_legend(override.aes = list(colour = NULL))) + scale_y_continuous(breaks=seq(0,30,5))

# Final plot
grid.arrange(g1, g2, ncol = 2, main = textGrob("VA Division Variability in Prescription Opioid Use Prior to TKA", gp=gpar(fontsize=24)))
```


## Jitter graph

Function to create "Five Thirty Eight" themed graphic
```{r}
fte_theme <- function() {
      
      # Generate the colors for the chart procedurally with RColorBrewer
      palette <- brewer.pal("Greys", n=9)
      color.background = palette[1]
      color.grid.major = palette[3]
      color.axis.text = palette[6]
      color.axis.title = palette[7]
      color.title = palette[9]
      
      # Begin construction of chart
      theme_bw(base_size=9) +
        
      # Set the entire chart region to a light gray color
      theme(panel.background=element_rect(fill=color.background, color=color.background)) +
      theme(plot.background=element_rect(fill=color.background, color=color.background)) +
      theme(panel.border=element_rect(color=color.background)) +
      
      # Format the grid
      theme(panel.grid.major=element_line(color=color.grid.major,size=.25)) +
      theme(panel.grid.minor=element_blank()) +
      theme(axis.ticks=element_blank()) +
      
      # Format the legend, but hide by default
      theme(legend.position="right") +
      theme(legend.background = element_rect(fill=color.background)) +
      theme(legend.text = element_text(size=20,color=color.axis.title)) +
      theme(legend.title = element_text(size=16)) +                      
      
      # Set title and axis labels, and format these and tick marks
      theme(plot.title=element_text(color=color.title, size=24, vjust=1.25)) +
      theme(axis.text.x=element_text(size=20,color=color.axis.text)) +
      theme(axis.text.y=element_text(size=20,color=color.axis.text)) +
      theme(axis.title.x=element_text(size=20,color=color.axis.title, vjust=0)) +
      theme(axis.title.y=element_text(size=20,color=color.axis.title, vjust=1.25)) +
      
      # Plot margins
      theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
    }
```
Jitter graph
```{r}
# Plot a and b row counts used for creating dataframes
n <- nrow(barData1)
l <- nrow(barData2)

# Plot a
jitData1 <- data.frame(Divisions = c(rep("Any Opioid\nUse", n), rep("Chronic\nUsers", n), rep(">100\nMEDs", n)))
jitData1$Percent <- c(barData1$Users, barData1$Chronic, barData1$HighMED)

b1<-ggplot(jitData1, aes(Divisions, Percent)) + 
  geom_jitter(alpha=I(3/4), aes(colour = Divisions), cex=3) +
  fte_theme() +
  theme(legend.position = "none") + 
  labs(title = "a)") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0))

# Plot b
jitData2 <- data.frame(Divisions = c(rep("Overall\nMedian MEDs", n), rep("Chronic User\nMedian MEDs", l)))
jitData2$Milligrams <- c(barData1$Median, barData2$Chron_Median)

b2<-ggplot(jitData2, aes(Divisions, Milligrams)) + 
  geom_jitter(alpha=I(3/4), aes(colour = Divisions), cex=3) +
  fte_theme() +
  labs(title = "b)") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0))

# Combine plots
grid.arrange(b1, b2, ncol = 2, main = textGrob("VA Division Variability in Opioid Prescribing Prior to TKA", gp=gpar(fontsize=24)))
```

## Division level results
```{r}
# 5th %ile and 95th %ile divisions for chronic use rate
quantile(barData1$Chronic, probs=c(.05, .95), na.rm=T)
```

## Map

Format statecode values
```{r}
statecodes$STATE_NAME <- as.character(statecodes$STATE_NAME)
statecodes$STATE <- formatC(statecodes$STATE, width = 2, flag = '0')
```
Format dataset location variables
```{r}
data$State <- as.character(data$State)
data$State <- gsub("INDIA0", "Indiana", data$State)
data$State <- gsub("SOUTH CAROLI0", "South Carolina", data$State)
data$State <- gsub("ARIZO0", "Arizona", data$State)
data$State <- gsub("NORTH CAROLI0", "North Carolina", data$State)
data$State <- gsub("LOUISIA0", "Louisiana", data$State)
data$State <- gsub("MONTA0", "Montana", data$State)
```
Function to format all state and county names appropriately
```{r}
simpleCap <- function(x) {
        if (x == "DISTRICT OF COLUMBIA") {
              return("District of Columbia")  
        } else {
                s <- tolower(x)
                s <- strsplit(s, " ")[[1]]
                paste(toupper(substring(s, 1,1)), substring(s, 2), sep="", collapse=" ")
        }
}
data$State <- sapply(data$State, simpleCap)
data$State <- gsub("*missing*", NA, data$State)
```
Function to convert state names to FIP codes
```{r}
recode <- function (x) {
        if (is.na(x) | x == "Armed Forces Af,eu,me,ca" | x == "Armed Forces Amer (exc Ca0da)") {
                y <- NA
        } else if (x == "Virginia") {
              y <- 51
        } else {
                y <- statecodes$STATE[grep(x, statecodes$STATE_NAME)]
        }
        y
}
data$STATEFP <- sapply(data$State, recode)
```
Format county values for mapping
```{r}
data$county <- as.character(data$county)
data$NAME <- sapply(data$county, simpleCap)
data$NAME <- gsub("0", NA, data$NAME)
```
Remove counties with less than 15 patients
```{r}
# Group data by state and county
datagrp <- group_by(data, STATEFP, NAME)

# Find the chronic use rate for all counties
chronicMapDF <- summarize(datagrp, Chronic_Use_Rate = mean(chronic), Count = n())

# Function to remove all counties with less than "min"
min <- 15
rmSmallSamp <- function() {
        for (i in 1:nrow(chronicMapDF)) {
                if (chronicMapDF$Count[i] < min) {
                        chronicMapDF$Chronic_Use_Rate[i] <- NA
                } else {next}
        }
        chronicMapDF
}
chronicMapDF <- rmSmallSamp() 
chronicMapDF$STATEFP <- as.factor(chronicMapDF$STATEFP)
chronicMapDF$NAME <- as.factor(chronicMapDF$NAME)
```
Combine US county map data with chronic opioid use summary data
```{r}
# Label all line data by associated county
counties@data$id <- rownames(counties@data)
# Combine line data with chronic opioid use data by county name
counties@data <- join(counties@data, chronicMapDF, by=c("STATEFP", "NAME"))
# Convert this combined dataset to a dataset ggplot can use
counties.df <- fortify(counties)
# I forgot what this does, but it's necessary to make it work
counties.df <- join(counties.df, counties@data, by="id")
# This is a world map, so subset the lat and long to zoom in on the US
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```
Create Map
```{r}
b <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = Chronic_Use_Rate), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", labels = percent, breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Chronic Opioid Use Prior to Total Knee Arthroplasty", fill = "") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .25))

#TIFF
ppi <- 300
tiff("EpiChronic.tiff", width = 6*ppi, height = 5*ppi, res = ppi)
b
dev.off()
```

## County-level results

```{r}
# 5th %ile and 95th %ile county for chronic use rate
quantile(chronicMapDF$Chronic_Use_Rate, probs=c(.05, .95), na.rm=T)
```