---
title: "OpiatesPredTKAMaps.Rmd"
author: "Jordan Starr"
date: "September 21, 2015"
output: html_document
---
##Load Packages & Data for Project Maps
```{r}
knitr::opts_chunk$set(cache=T)
library("ggplot2")
library("ggmap")
library("plyr")
library("rgdal")
library("scales")
library("Cairo")
library("maptools")
library("rgeos")
library("dplyr")
library("grid")
data <- read.csv("WorkDf5.csv", sep=";")
statecodes <- read.delim("./state.txt", sep="|", header=T)
statecodes$STATE_NAME <- as.character(statecodes$STATE_NAME)
statecodes$STATE <- formatC(statecodes$STATE, width = 2, flag = '0')
```

##Clean Data

Remove unnecessary columns
```{r}
data <- data[,c(9,10,19,20,39,47)]
```
Create MEDperDay variables and remove extreme outliers
```{r}
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
ggplot(data=data, aes(data$MEDperMonth)) + geom_histogram() + coord_cartesian(ylim=c(0,5)) + scale_x_continuous(breaks=seq(50000,150000,10000))
data$MEDperMonth <- ifelse(data$MEDperMonth > 80000, NA, data$MEDperMonth)
data$MEDperDay <- data$MEDperMonth/(365/12)
data$highMED <- ifelse(data$MEDperDay > 100, 1, 0)
data$user <- ifelse(data$MEDperDay > 0, 1, 0)
```
Create chronic variable
```{r}
data$chronic <- ifelse(data$TimeBefore > 2 & data$MEDperDay >= 2, 1, 0)
# For Chronic User Median MED map only
# data <- data[data$chronic == 1,]
```
Format state values for mapping
```{r}
unique(data$State)
data$State <- as.character(data$State)
data$State <- gsub("INDIA0", "Indiana", data$State)
data$State <- gsub("SOUTH CAROLI0", "South Carolina", data$State)
data$State <- gsub("ARIZO0", "Arizona", data$State)
data$State <- gsub("NORTH CAROLI0", "North Carolina", data$State)
data$State <- gsub("LOUISIA0", "Louisiana", data$State)
data$State <- gsub("MONTA0", "Montana", data$State)

#Function to format all state and county names
simpleCap <- function(x) {
        if (x == "DISTRICT OF COLUMBIA") {
              return("District of Columbia")  
        } else {
                s <- tolower(x)
                s <- strsplit(s, " ")[[1]]
                paste(toupper(substring(s, 1,1)), substring(s, 2), sep="", collapse=" ")
        }
}
## For Chronic user median MED map
## data <- data[!is.na(data$State),]
data$State <- sapply(data$State, simpleCap)
data$State <- gsub("*missing*", NA, data$State)

#Function to convert state names to FIP codes
recode <- function (x) {
        if (is.na(x) | x == "Armed Forces Af,eu,me,ca" | x == "Armed Forces Amer (exc Ca0da)") {
                y <- NA
        } else if (x == "Virginia") {
              y <- 51
        } else {
                y <- statecodes$STATE[grep(x, statecodes$STATE_NAME)]
        }
        y
}
data$STATEFP <- sapply(data$State, recode)
```
Format county values for mapping
```{r}
data$county <- as.character(data$county)
data$NAME <- sapply(data$county, simpleCap)
data$NAME <- gsub("0", NA, data$NAME)
```
Remove unnecessary columns again and remove NAs
```{r}
data <- data[,c(10:13)]
data <- na.omit(data)
```
Group and summarize data by state and county, then remove counties with less than 15 patients.
```{r}
datagrp <- group_by(data, STATEFP, NAME)

#Revision Rate Map Data
revMapDF <- summarize(datagrp, Revision_Rate = mean(RevLogical), Count = n())
rmSmallSamp1 <- function() {
        for (i in 1:nrow(revMapDF)) {
                if (revMapDF$Count[i] < 15) {
                        revMapDF$Revision_Rate[i] <- NA
                } else {next}
        }
        revMapDF
}
revMapDF <- rmSmallSamp1() 
revMapDF$NAME <- as.factor(revMapDF$NAME)
revMapDF$STATEFP <- as.factor(revMapDF$STATEFP)

#MED/Day Map Data 
MEDMapDF <- summarize(datagrp, Median_MED_per_Day = median(MEDperDay), Count = n())
rmSmallSamp2 <- function() {
        for (i in 1:nrow(MEDMapDF)) {
                if (MEDMapDF$Count[i] < 15) {
                        MEDMapDF$Median_MED_per_Day[i] <- NA
                } else {next}
        }
        MEDMapDF
}
MEDMapDF <- rmSmallSamp2() 
MEDMapDF$STATEFP <- as.factor(MEDMapDF$STATEFP)
MEDMapDF$NAME <- as.factor(MEDMapDF$NAME)

#Chronic Opiates Map Data
chronicMapDF <- summarize(datagrp, Chronic_Use_Rate = mean(chronic), Count = n())
rmSmallSamp3 <- function() {
        for (i in 1:nrow(chronicMapDF)) {
                if (chronicMapDF$Count[i] < 15) {
                        chronicMapDF$Chronic_Use_Rate[i] <- NA
                } else {next}
        }
        chronicMapDF
}
chronicMapDF <- rmSmallSamp3() 
chronicMapDF$STATEFP <- as.factor(chronicMapDF$STATEFP)
chronicMapDF$NAME <- as.factor(chronicMapDF$NAME)

#TypeBefore Map Data
typesMapDF <- summarize(datagrp, Type_Before = mean(TypeBefore), Count = n())
rmSmallSamp4 <- function() {
        for (i in 1:nrow(typesMapDF)) {
                if (typesMapDF$Count[i] < 15) {
                        typesMapDF$Type_Before[i] <- NA
                } else {next}
        }
        typesMapDF
}
typesMapDF <- rmSmallSamp4() 
typesMapDF$STATEFP <- as.factor(typesMapDF$STATEFP)
typesMapDF$NAME <- as.factor(typesMapDF$NAME)

#highMED data
highMEDMapDF <- summarize(datagrp, High_MED_Rate = mean(highMED), Count = n())
rmSmallSamp5 <- function() {
        for (i in 1:nrow(highMEDMapDF)) {
                if (highMEDMapDF$Count[i] < 15) {
                        highMEDMapDF$High_MED_Rate[i] <- NA
                } else {next}
        }
        highMEDMapDF
}
highMEDMapDF <- rmSmallSamp5() 
highMEDMapDF$STATEFP <- as.factor(highMEDMapDF$STATEFP)
highMEDMapDF$NAME <- as.factor(highMEDMapDF$NAME)

#user data
userMapDF <- summarize(datagrp, User_Rate = mean(user), Count = n())
rmSmallSamp6 <- function() {
        for (i in 1:nrow(userMapDF)) {
                if (userMapDF$Count[i] < 15) {
                        userMapDF$User_Rate[i] <- NA
                } else {next}
        }
        userMapDF
}
userMapDF <- rmSmallSamp6() 
userMapDF$STATEFP <- as.factor(userMapDF$STATEFP)
userMapDF$NAME <- as.factor(userMapDF$NAME)

#Chronic user median MED map
chronMedMEDMapDF <- summarize(datagrp, Median_MED = median(MEDperDay), Count = n())
rmSmallSamp7 <- function() {
        for (i in 1:nrow(chronMedMEDMapDF)) {
                if (chronMedMEDMapDF$Count[i] < 10) {
                        chronMedMEDMapDF$Median_MED[i] <- NA
                } else {next}
        }
        chronMedMEDMapDF
}
chronMedMEDMapDF <- rmSmallSamp7() 
chronMedMEDMapDF$STATEFP <- as.factor(chronMedMEDMapDF$STATEFP)
chronMedMEDMapDF$NAME <- as.factor(chronMedMEDMapDF$NAME)
```
Load map data and combine with revision data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, revMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```
Create Revision Map
```{r}
j <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = Revision_Rate), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", labels = percent, breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "TKA Revision Rate by County \n Number of Revisions/ All TKAs", fill = "") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))

#PNG
ppi <- 300
png("TKA Revision.png", width = 6*ppi, height = 6*ppi, res = ppi)
j
dev.off()

#PDF
pdf("TKA Revision.pdf")
j
dev.off()
```
Reload map data and combine with MED/Day data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, MEDMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```
Create MED/Day Map
```{r}
p <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = Median_MED_per_Day), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Median Preoperative Daily Oral Morphine Equivalent\nUsage per Day by County", fill = "Milligrams") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))

#PNG
ppi <- 300
png("MED.png", width = 6*ppi, height = 6*ppi, res = ppi)
p
dev.off()

#PDF
pdf("MED.pdf")
p
dev.off()
```

Reload map data and combine with chronic opiate use data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, chronicMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```

Create Chronic Opiate Use Rate Map
```{r}
b <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = Chronic_Use_Rate), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", labels = percent, breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Rate of Patients Using Opiates Chronically \n Preoperatively by County", fill = "") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))

#PNG
ppi <- 300
png("Chronic.png", width = 6*ppi, height = 6*ppi, res = ppi)
b
dev.off()

#PDF
pdf("Chronic.pdf")
b
dev.off()
```

Reload map data and combine with typeBefore data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, typesMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```

Create typeBefore Map
```{r}
c <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = Type_Before), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Mean Number of Unique Opiates Prescribed \n Preoperatively by County", fill = "") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))

#PNG
ppi <- 300
png("TypeBefore.png", width = 6*ppi, height = 6*ppi, res = ppi)
c
dev.off()

#PDF
pdf("TypeBefore.pdf")
c
dev.off()
```
Reload map data and combine with highMED data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, highMEDMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```

Create High MED Map
```{r}
h <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = High_MED_Rate), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", labels = percent, breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Rate of Patients Using Greater than 100 Daily Oral Morphine Equivalents\nin Milligrams Preoperatively by County", fill = "") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))
```

Reload map data and combine with user data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, userMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```

Create User Map
```{r}
h <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = User_Rate), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", labels = percent, breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Rates of Patients Using Opioids Preoperatively by County", fill = "") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))
```
Reload map data and combine with chronic user median MED data
```{r}
counties <- readOGR(dsn = "./cb_2014_us_county_500k", layer = "cb_2014_us_county_500k")
counties@data$id <- rownames(counties@data)
counties@data <- join(counties@data, chronMedMEDMapDF, by=c("STATEFP", "NAME"))
counties.df <- fortify(counties)
counties.df <- join(counties.df, counties@data, by="id")
counties.df2 <- counties.df[counties.df$long < -60 & counties.df$long > -130 & counties.df$lat > 25 & counties.df$lat < 50,]
```

Create Chronic User Median MED Map
```{r}
h <- ggplot() + geom_polygon(data = counties.df2, aes(x = long, y = lat, group = group, fill = Median_MED), color = "black", size = 0.05)+ coord_map() + scale_fill_distiller(palette = "YlOrRd", na.value = "white", breaks = pretty_breaks(n = 10)) + guides(fill = guide_legend(reverse = TRUE)) + theme_nothing(legend = TRUE) + labs(title = "Median Daily Oral Morphine Equivalent Usage\nAmong Chronic Opiate Users by County", fill = "Milligrams") + theme(plot.title = element_text(size=20, family = "Times"),legend.key.height = unit(.35, "cm"),legend.position = c(.9, .3))
```