---
title: "PredTKARev2"
author: "Jordan Starr"
date: "November 18, 2015"
output: html_document
---
##Load Data & Packages
```{r cache=TRUE}
knitr::opts_chunk$set(cache=T)
library("dplyr")
library("caret")
library("ROCR")
library("scales")
library("randomForest")
set.seed(12)
alldata <- read.csv("WorkDf5.csv", sep=";")
```

#Clean Data
```{r}
data <- alldata[,c(5:8,19,26:32,39,47,49)]
data$Diabetes <- as.factor(data$Diabetes)
data$CKD <- as.factor(data$CKD)
data$HTN <- as.factor(data$HTN)
data$CHF <- as.factor(data$CHF)
data$OSA <- as.factor(data$OSA)
data$PTSD <- as.factor(data$PTSD)
data$Tobbaco <- as.factor(data$Tobbaco)
data$ht <- ifelse(data$ht < 48 | data$ht > 86, NA, data$ht)
data$wt <- ifelse(data$wt < 80 | data$wt > 600, NA, data$wt)
data <- mutate(data, BMI = (wt/2.2)/(ht*2.54/100)^2)
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
data$MEDperMonth <- ifelse(data$MEDperMonth > 80000, NA, data$MEDperMonth)
```

#5 year f/u
```{r}
fu5 <- function() {
        x <- vector()
        for (i in 1:nrow(data)) {
                if (data$RevLogical[i] == 0 & data$t2rev[i] < 1825) {
                        x <- append(x, NA)
                }
                else if (data$RevLogical[i] == 1 & data$t2rev[i] > 1825) {
                        x <- append(x, 0)
                }
                else {
                        x <- append(x, data$RevLogical[i])   
                }
        }
        return(x)
}
data$fu5 <- fu5()
data$fu5 <- factor(data$fu5, levels = c(0,1),labels = c("no", "yes"))
data <- data[,c(-3:-5,-13:-15)]
data <- na.omit(data)
```

Training, testing, and validation datasets:
```{r}
inTrain <- createDataPartition(y=data$fu5, p = .6, list=FALSE)
training <- data[inTrain, ]
testingboth <- data[-inTrain, ]
inTrain2 <- createDataPartition(y=testingboth$fu5, p = .5, list =F)
testing <- testingboth[inTrain2,]
validation <- testingboth[-inTrain2,]
```

GLM with 10-fold cross validation
```{r}
ctrl <- trainControl(method="cv", number=10, classProbs = TRUE)
glmfit <- train(fu5 ~ ., data = training, method="glm", trControl = ctrl)
pred1 <- predict(glmfit, testing, type="prob")
x <- prediction(pred1[,2],testing$fu5)
y <- performance(x, measure = "tpr", x.measure = "fpr")
plot(y)
abline(a=0, b= 1)
z <- performance(x, measure = "auc")
p <- unlist(z@y.values)
n <- nrow(training)
s = sqrt(p*(1-p)/n)
CI <- c(p-1.96*s, p+1.96*s)
p #AUC
CI #AUC 95% CI
```

Random Forest
```{r}
rffit <- randomForest(fu5 ~ ., data=training, importance=TRUE)
pred2 <- predict(rffit, testing, type="prob")
x <- prediction(pred2[,2],testing$fu5)
y <- performance(x, measure = "tpr", x.measure = "fpr")
plot(y)
abline(a=0, b= 1)
z <- performance(x, measure = "auc")
p <- unlist(z@y.values)
n <- nrow(training)
s = sqrt(p*(1-p)/n)
CI <- c(p-1.96*s, p+1.96*s)
p #AUC
CI #AUC 95% CI
```

Rpart
```{r}
rpartfit <- train(fu5 ~ ., data = training, method="rpart")
pred3 <- predict(rpartfit, testing, type="prob")
x <- prediction(pred3[,2],testing$fu5)
y <- performance(x, measure = "tpr", x.measure = "fpr")
plot(y)
abline(a=0, b= 1)
z <- performance(x, measure = "auc")
p <- unlist(z@y.values)
n <- nrow(training)
s = sqrt(p*(1-p)/n)
CI <- c(p-1.96*s, p+1.96*s)
p #AUC
CI #AUC 95% CI
```

Boosting
```{r}
gbmfit <- train(fu5 ~ ., data = training, method="gbm", verbose = F)
detach("package:dplyr", unload=T) # Need to reload dplyr after gbm to avoid errors
library(dplyr)
pred4 <- predict(gbmfit, testing, type="prob")
x <- prediction(pred4[,2],testing$fu5)
y <- performance(x, measure = "tpr", x.measure = "fpr")
plot(y)
abline(a=0, b= 1)
z <- performance(x, measure = "auc")
p <- unlist(z@y.values)
n <- nrow(training)
s = sqrt(p*(1-p)/n)
CI <- c(p-1.96*s, p+1.96*s)
p #AUC
CI #AUC 95% CI
```

Glmnet with 10-fold cross validation
```{r}
ctrl <- trainControl(method="cv", number=10, classProbs = TRUE)
glmnetfit <- train(fu5 ~ ., data = training, method="glmnet", trControl = ctrl)
pred5 <- predict(glmnetfit, testing, type = "prob")
x <- prediction(pred5[,2],testing$fu5)
y <- performance(x, measure = "tpr", x.measure = "fpr")
plot(y)
abline(a=0, b= 1)
z <- performance(x, measure = "auc")
p <- unlist(z@y.values)
n <- nrow(training)
s = sqrt(p*(1-p)/n)
CI <- c(p-1.96*s, p+1.96*s)
p #AUC
CI #AUC 95% CI
```

Combined Model
```{r}
predDF <- data.frame(pred1 = pred1[,2], pred2 = pred2[,2], pred3 = pred3[,2], pred4 = pred4[,2], pred5 = pred5[,2], fu5 = testing$fu5)
combModFit <- train(fu5 ~ ., data = predDF, method = "gam")
pred1V <- predict(glmfit, validation, type = "prob")
pred2V <- predict(rffit, validation, type = "prob")
pred3V <- predict(rpartfit, validation, type = "prob")
pred4V <- predict(gbmfit, validation, type = "prob")
pred5V <- predict(glmnetfit, validation, type = "prob")
predVDF <- data.frame(pred1 = pred1V[,2], pred2 = pred2V[,2], pred3 = pred3V[,2], pred4 = pred4V[,2], pred5 = pred5V[,2])
combPredV <- predict(combModFit, predVDF, type = "prob")
x <- prediction(combPredV[,1],validation$fu5)
y <- performance(x, measure = "tpr", x.measure = "fpr")
plot(y)
abline(a=0, b= 1)
z <- performance(x, measure = "auc")
p <- unlist(z@y.values)
n <- nrow(training)
s = sqrt(p*(1-p)/n)
CI <- c(p-1.96*s, p+1.96*s)
p #AUC
CI #AUC 95% CI