Load libraries and minimally clean data
```{r}
knitr::opts_chunk$set(cache=T)
library("ggplot2")
library("dplyr")
library(grid)
library(gridExtra)
data <- read.csv("WorkDf5.csv", sep=";")
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
data$MEDperMonth <- ifelse(data$MEDperMonth > 80000, NA, data$MEDperMonth)
data$MEDperDay <- data$MEDperMonth/(365/12)
data$realChronic <- ifelse(data$TimeBefore > 2 & data$MEDperDay >= 2, 1, 0)
data$highMED <- ifelse(data$MEDperDay > 100, 1, 0)
data$user <- ifelse(data$MEDperDay > 0,1,0)
data$ht <- ifelse(data$ht < 48 | data$ht > 86, NA, data$ht)
data$wt <- ifelse(data$wt < 80 | data$wt > 600, NA, data$wt)
data <- mutate(data, BMI = (wt/2.2)/(ht*2.54/100)^2)
datagrp <- group_by(data, TKAYear)
data$TKAYear <- as.factor(data$TKAYear)
```
#Create Table 1
```{r}
mean(data$Age)
sd(data$Age)
mean(data$BMI, na.rm=T)
sd(data$BMI, na.rm=T)
nrow(data[data$gender == "M",])/ nrow(data)
sum(data$CKD, na.rm=T)/ nrow(data)
sum(data$Diabetes, na.rm=T)/ nrow(data)
sum(data$HTN, na.rm=T)/ nrow(data)
sum(data$OSA, na.rm=T)/ nrow(data)
sum(data$PTSD, na.rm=T)/ nrow(data)
data2 <- filter(data, MEDperDay == 0)
mean(data2$Age)
sd(data2$Age)
mean(data2$BMI, na.rm=T)
sd(data2$BMI, na.rm=T)
nrow(data2[data2$gender == "M",])/ nrow(data2)
sum(data2$CKD, na.rm=T)/ nrow(data2)
sum(data2$Diabetes, na.rm=T)/ nrow(data2)
sum(data2$HTN, na.rm=T)/ nrow(data2)
sum(data2$OSA, na.rm=T)/ nrow(data2)
sum(data2$PTSD, na.rm=T)/ nrow(data2)
data2 <- filter(data, MEDperDay > 0 & realChronic == 0)
mean(data2$Age)
sd(data2$Age)
mean(data2$BMI, na.rm=T)
sd(data2$BMI, na.rm=T)
nrow(data2[data2$gender == "M",])/ nrow(data2)
sum(data2$CKD, na.rm=T)/ nrow(data2)
sum(data2$Diabetes, na.rm=T)/ nrow(data2)
sum(data2$HTN, na.rm=T)/ nrow(data2)
sum(data2$OSA, na.rm=T)/ nrow(data2)
sum(data2$PTSD, na.rm=T)/ nrow(data2)
data2 <- filter(data, realChronic == 1)
mean(data2$Age)
sd(data2$Age)
mean(data2$BMI, na.rm=T)
sd(data2$BMI, na.rm=T)
nrow(data2[data2$gender == "M",])/ nrow(data2)
sum(data2$CKD, na.rm=T)/ nrow(data2)
sum(data2$Diabetes, na.rm=T)/ nrow(data2)
sum(data2$HTN, na.rm=T)/ nrow(data2)
sum(data2$OSA, na.rm=T)/ nrow(data2)
sum(data2$PTSD, na.rm=T)/ nrow(data2)

#P values
prop.test(x=c(.955*14077,.937*5122,.936*14450), n=c(14077, 5122, 14450))
prop.test(x=c(.083*14077,.096*5122,.082*14450), n=c(14077, 5122, 14450))
prop.test(x=c(.313*14077,.320*5122,.320*14450), n=c(14077, 5122, 14450))
prop.test(x=c(.665*14077,.675*5122,.653*14450), n=c(14077, 5122, 14450))
prop.test(x=c(.051*14077,.060*5122,.066*14450), n=c(14077, 5122, 14450))
prop.test(x=c(.127*14077,.157*5122,.201*14450), n=c(14077, 5122, 14450))

data$usergrp <- ifelse(data$user == 0, 0, 1)
data$usergrp <- ifelse(data$realChronic == 1, 2, data$usergrp)
fit1 <- aov(data$Age ~ data$usergrp)
summary(fit1)
fit2 <- aov(data$BMI ~ data$usergrp)
summary(fit2)
```
#Create Table 2
```{r}
table1 <- summarise(datagrp, count = n(), users = sum(user, na.rm=T)/n(), medMED = median(MEDperDay, na.rm=T), IQR25 = quantile(MEDperDay, na.rm=T)[2], IQR75 = quantile(MEDperDay, na.rm=T)[4], chronic = sum(realChronic, na.rm=T)/n(), hundred = sum(highMED, na.rm=T)/n())
table1

datagrp2 <- filter(datagrp, realChronic == 1)
table2 <- summarise(datagrp2, chron_med_MED = median(MEDperDay, na.rm=T), IQR25 = quantile(MEDperDay, na.rm=T)[2], IQR75 = quantile(MEDperDay, na.rm=T)[4] )
table2

#P values
year <- c(2006, 2007, 2008, 2009, 2010, 2011)
overall <- c(5296, 5103, 5551, 5751, 6027, 5914)
users <- c(.598,.570,.571,.584,.581,.583)
median <- c(1.47,1.23,1.21,1.41,1.32,1.41)
chronic <- c(.434, .427,.422,.433,.432,.43)
chronMed <- c(10.06,10.99,10.89,11.10,12.1,13.15)
highMED <- c(2.1,2.2,2.4,1.8,1.9,2.1)
fit1 <- lm(overall ~ year)
fit2 <- lm(users ~ year)
fit3 <- lm(median ~ year)
fit4 <- lm(chronic ~ year)
fit5 <- lm(chronMed ~ year)
fit6 <- lm(highMED ~ year)
summary(fit1)
summary(fit2)
summary(fit3)
summary(fit4)
summary(fit5)
summary(fit6)
```
Bar Graph
```{r}
datagrp3 <- group_by(data, DIVISION)
barData1 <- summarize(datagrp3, N = n(), Users = sum(user, na.rm=T)/n(), Chronic = sum(realChronic, na.rm=T)/n(), HighMED = sum(highMED, na.rm=T)/n(), Median = median(MEDperDay, na.rm=T))
barData1 <- barData1[barData1$N > 49,]
datagrp4 <- group_by(data[data$realChronic == 1,], DIVISION)
barData2 <- summarize(datagrp4, N = n(), Chron_Median = median(MEDperDay, na.rm=T))
barData2 <- barData2[barData2$N > 24,]

barDF1 <- data.frame(Outcome = c(rep("Opioid Users", 2), rep("Chronic Opioid\nUsers", 2), rep(">100 MEs\nper Day", 2)), Group = c(rep(c("5th %", "95th %"),3)))
barDF1$percent <- c(quantile(barData1$Users, c(.05,.95)), quantile(barData1$Chronic, c(.05,.95)), quantile(barData1$HighMED, c(.05,.95)))

barDF2 <- data.frame(Outcome = c(rep("Overall\nMedian MEs", 2), rep("Chronic User\nMedian MEs", 2)), Group = c(rep(c("5th %", "95th %"),2)))
barDF2$median <- c(quantile(barData1$Median, c(.05,.95)), quantile(barData2$Chron_Median, c(.05,.95)))

barDF1$Outcome <- factor(barDF1$Outcome, levels=barDF1$Outcome)
barDF2$Outcome <- factor(barDF2$Outcome, levels=barDF2$Outcome)
g1 <- ggplot(barDF1, aes(Outcome, percent, fill=Group)) + geom_bar(position="dodge", stat="identity", color = "black") + theme_bw(base_size=18) + scale_fill_brewer() + labs(x="", y ="Rate") + theme(legend.title=element_blank()) + coord_cartesian(ylim=c(0,.7)) + scale_y_continuous(breaks=seq(0,.7,.1)) + guides(fill=FALSE) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

g2 <- ggplot(barDF2, aes(Outcome, median, fill=Group)) + geom_bar(position="dodge", stat="identity", color = "black") + theme_bw(base_size=18) + coord_cartesian(ylim=c(0,18)) + scale_fill_brewer() + labs(x="", y ="Milligrams") + theme(legend.title=element_blank()) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + guides(fill = guide_legend(override.aes = list(colour = NULL))) + scale_y_continuous(breaks=seq(0,18,3))

grid.arrange(g1, g2, ncol = 2, main = textGrob("Variability in Prescription Opioid Use by VA Medical Centers", gp=gpar(fontsize=24)))
```

State analysis
```{r}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

datagrp3 <- group_by(data, DIVISION)
barData1 <- summarize(datagrp3, State = Mode(State), N = n(), Users = sum(user, na.rm=T)/n(), Chronic = sum(realChronic, na.rm=T)/n(), HighMED = sum(highMED, na.rm=T)/n(), Median = median(MEDperDay, na.rm=T))
barData1 <- barData1[barData1$N > 49,]
datagrp4 <- group_by(data[data$realChronic == 1,], DIVISION)
barData2 <- summarize(datagrp4, N = n(), State = Mode(State), Chron_Median = median(MEDperDay, na.rm=T))
barData2 <- barData2[barData2$N > 24,]

View(arrange(barData1, Users))
View(arrange(barData1, Chronic))
View(arrange(barData1, HighMED))
View(arrange(barData1, Median))
View(arrange(barData2, Chron_Median))
```
Graph: MEDperMonth by year 
```{r}
#Mean
meanMEDxYear <- summarise(datagrp, Mean_MED = mean(MEDperMonth, na.rm=T))
g <- ggplot(meanMEDxYear, aes(TKAYear, Mean_MED)) + geom_point() + geom_smooth(method = "lm", se = F) + labs(title = "Mean Preoperative Morphine Equivalents \n per Month by Year", x = "Year", y = "Mean Morphine Equivalents per Month")

#Median
medianMEDxYear <- summarise(datagrp, MEDperMonth = median(MEDperMonth, na.rm=T))
p <- ggplot(medianMEDxYear, aes(TKAYear, Median_MED)) + geom_point() + geom_smooth(method = "lm", se = F) + labs(title = "Median Preoperative Morphine Equivalents \n per Month by Year", x = "Year", y = "Median Morphine Equivalents per Month")

#Median graph with quartiles
p <- ggplot(data, aes(TKAYear, MEDperMonth)) + geom_boxplot(outlier.shape = NA, fill="aliceblue") + coord_cartesian(ylim = c(0,875))+geom_smooth(method="lm", aes(group=1))+ labs(title = "Preoperative Opiate Use by Year", x = "Year", y = "Distribution of Morphine Equivalents per Month")
```

Graph: TypeBefore by year
```{r}
#Mean
meanTBxYear <- summarise(datagrp, Mean_Types = mean(TypeBefore, na.rm=T))
a <- ggplot(meanTBxYear, aes(TKAYear, Mean_Types)) + geom_point() + geom_smooth(method = "lm", se = F) + labs(title = "Mean Number of Unique Opiates Prescribed \n Preoperatively by Year", x = "Year", y = "Mean Unique Opiates Prescribed")

#Median
medianTBxYear <- summarise(datagrp, Median_Types = median(TypeBefore, na.rm=T))
b <- ggplot(medianTBxYear, aes(TKAYear, Median_Types)) + geom_point() + geom_smooth(method = "lm", se = F) + labs(title = "Median Number of Unique Opiates Prescribed \n Preoperatively by Year", x = "Year", y = "Median Unique Opiates Prescribed")
```

Create graphs
```{r}
ppi <- 300
png("MeanMEDxYear.png", width = 6*ppi, height = 6*ppi, res = ppi)
g
dev.off()
pdf("MeanMEDxYear.pdf")
g
dev.off()
png("MedianMEDxYear.png", width = 6*ppi, height = 6*ppi, res = ppi)
p
dev.off()
pdf("MedianMEDxYear.pdf")
p
dev.off()
png("MeanTypeBeforexYear.png", width = 6*ppi, height = 6*ppi, res = ppi)
a
dev.off()
pdf("MeanTypeBeforexYear.pdf")
a
dev.off()
png("MedianTypeBeforexYear.png", width = 6*ppi, height = 6*ppi, res = ppi)
b
dev.off()
pdf("MedianTypeBeforexYear.pdf")
b
dev.off()
```

Age histogram, grouped by chronic opioid use
```{r}
data$chronic <- as.factor(data$chronic)
data <- rename(data, Chronic = chronic)
data$Chronic <- ifelse(data$Chronic == 1, "Yes", "No")
z <- ggplot(data, aes(x=Age, fill = Chronic)) + geom_histogram() + labs(title = "Age Distribution Grouped by Chronic \n Opiate Use", x = "Age", y = "Count")

#Save
ppi <- 300
png("AgexChronic.png", width = 6*ppi, height = 6*ppi, res = ppi)
z
dev.off()
pdf("AgexChronic.pdf")
z
dev.off()
```

Revision free survival curve separated by chronic opioid use with CIs and p-value
```{r}
library("survival")
library("rms")
data$chronic <- ifelse(data$chronic == 1, "Chronic", "Not Chronic")
data$chronic <- as.factor(data$chronic)
data$t2rev <- ifelse(data$t2rev < 0, NA, data$t2rev)
data2 <- na.omit(data[,c(39,48,49)])
data2$RevLogical <- as.logical(ifelse(data2$RevLogical == 0, F, T))
fit <- survdiff(Surv(data2$t2rev, data2$RevLogical) ~ data2$chronic)
fit #gives p-value
objNpsurv <- npsurv(formula = Surv(t2rev, RevLogical == 1) ~ chronic, data = data2)
ppi <- 300
png("Survival.png", width = 6*ppi, height = 6*ppi, res = ppi)
survplot(objNpsurv, ylim = c(.9,1), xlab = "Days", ylab = "Percent of TKAs Unrevised", conf.type="log-log", levels.only=T, col.fill =c(7,6), legend.pos="right")
title(main = "Survival Curve of TKA Revision by Opiate Use")
text(2100, .98, "P = 1.71 x10^-11")
dev.off()
pdf("Survival.pdf")
survplot(objNpsurv, ylim = c(.9,1), xlab = "Days", ylab = "Percent of TKAs Unrevised", conf.type="log-log", levels.only=T, col.fill =c(7,6), legend.pos="right")
title(main = "Survival Curve of TKA Revision by Opiate Use")
text(2100, .98, "P = 1.71 x10^-11")
dev.off()
```

Surface plot: x = MED/Month, y = TimeBefore, z = count
```{r}
#Prepare data
library(MASS)
library(RColorBrewer)
library(hexbin)
rf <- colorRampPalette(rev(brewer.pal(11,'Spectral')))
r <- rf(32)
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
data$MEDperMonth <- ifelse(data$MEDperMonth > 80000, NA, data$MEDperMonth) #Remove outliers
data <- mutate(data, MEDperDay = MEDperMonth/(365/12))
data2 <- na.omit(data[,c(51,19)])
data3 <- data2[data2$TimeBefore > 2 & data2$MEDperDay >=2,] #Only graph chronic opiate users

#ggplot version of graph
p <- ggplot(data3, aes(MEDperDay,TimeBefore))
h3 <- p + stat_bin2d(bins=50) + scale_fill_gradientn(colours=r) + labs(title="Density Plot of Preop Morphine Equivalent \n Use in Chronic Users", x="Morphine Equivalents/ Day", y="Months")
h3

#hexbin version of graph
hexbinplot(TimeBefore ~ MEDperDay, data=data3, colramp=rf, trans=log, inv=exp)

#KDE2D version of graph
k <- kde2d(data3[,1], data3[,2], n=200, lims=c(0,75,3,30))
#Pretty
image(k, col=r, ylim=c(0,30))
#Less pretty
filled.contour(k,color.palette=colorRampPalette(c('white','blue','yellow','red','darkred')), ylim=c(0,25), xlim=c(0,50))

#Code to save KDE2D version of graph
ppi <- 300
png("Density.png", width = 6*ppi, height = 6*ppi, res = ppi)
image(k, col=r, ylim=c(0,30))
title(main="(b)", xlab = "Morphine Equivalents / Day", ylab = "Months")
dev.off()
pdf("Density.pdf")
image(k, col=r, ylim=c(0,30))
title(main="(b)", xlab = "Morphine Equivalents / Day", ylab = "Months")
dev.off()

#Add Histograms
hist(data3$MEDperDay, breaks = seq(2,2500,2), freq = F, xlim = c(2,73), main = "(a)", xlab = "Morphine Equivalents/ Day", ylab = "Percent of Chronic Opiate Users", xaxt="n", col = "grey", ylim=c(0,.1))
axis(1, at=c(0,10,20,30,40,50,60,70))

A <- hist(data3$TimeBefore, breaks = seq(3,71,2), freq = F, xlim = c(0,29))
plot(NULL, type = "n", xlim = c(0, max(A$density)), ylim = c(0,30),main = "(c)", ylab = "Months", xlab = "Percent of Chronic Opiate Users")
rect(0, A$breaks[1:(length(A$breaks) - 1)], A$density, A$breaks[2:length(A$breaks)], col = "grey")

#Images together
layout(matrix(c(3,4,1,2),2,2,byrow=T))
image(k, col=r, ylim=c(0,30))
title(main="(b)", xlab = "Morphine Equivalents / Day", ylab = "Months")

plot(NULL, type = "n", xlim = c(0, max(A$density)), ylim = c(0,30),main = "(c)", ylab = "Months", xlab = "Percent of Chronic Opiate Users")
rect(0, A$breaks[1:(length(A$breaks) - 1)], A$density, A$breaks[2:length(A$breaks)], col = "grey")

hist(data3$MEDperDay, breaks = seq(2,2500,2), freq = F, xlim = c(2,73), main = "(a)", xlab = "Morphine Equivalents/ Day", ylab = "Percent of Chronic Opiate Users", xaxt="n", col = "grey", ylim=c(0,.1))
axis(1, at=c(0,10,20,30,40,50,60,70))
```

Devision scatterplot 1
```{r}
library("ggplot2")
library("dplyr")
data <- read.csv("WorkDf5.csv", sep=";")
data <- data[,c(11,19,39,47)]
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
data$MEDperMonth <- ifelse(data$MEDperMonth > 80000, NA, data$MEDperMonth)
data <- mutate(data, MEDperDay = MEDperMonth/(365/12))
data <- mutate(data, Chronic = ifelse(TimeBefore > 2 & MEDperDay >= 2, 1, 0))
data <- na.omit(data[,c(1,3,6,7)])
datagrp <- group_by(data, DIVISION)
points <- summarize(datagrp, Count = n(), Rev_Rate = sum(RevLogical)/n(), Med_MED = median(MEDperDay), Chron_Rate = sum(Chronic)/n())
points <- points[points$Count > 49,]
meanMED <- mean(points$Med_MED)
sdMED <- sd(points$Med_MED)
vert <- meanMED + (2 * sdMED)  
meanRev <- mean(points$Rev_Rate) 
sdRev <- sd(points$Rev_Rate)
hor <- meanRev + (2 * sdRev)  

#Plot
p <- ggplot(points, aes(Med_MED, Rev_Rate))
p + geom_hline(aes(yintercept=hor), color = "blue", linetype = 2) + geom_vline(aes(xintercept=vert), color = "blue", linetype = 2) + geom_point(aes(colour = Chron_Rate), size = 5) + scale_colour_gradient(high = "red", breaks =c(.1,.2,.3,.4,.5), guide = guide_legend(title = "Chronic Opiate \n Use Rate", reverse = T, keywidth = 3)) + theme_bw() + labs(title = "Opiate Use and TKA Revision for Each VAMC Division", x = "Median Daily Morphine Equivalents", y = "TKA Revision Rate") + annotate("text", x = (vert+.55), y = (hor+.002), colour = "blue", label = "Mean + 2 SD") 
```

Devision scatterplot 2
```{r}
library("ggplot2")
library("dplyr")
data <- read.csv("WorkDf5.csv", sep=";")
data <- data[,c(11,19,47)]
data$MEDperMonth <- ifelse(data$TimeBefore==0, 0, data$med_load/data$TimeBefore)
data$MEDperMonth <- ifelse(data$MEDperMonth > 80000, NA, data$MEDperMonth)
data <- mutate(data, MEDperDay = MEDperMonth/(365/12))
data <- mutate(data, Chronic = ifelse(TimeBefore > 2 & MEDperDay >= 2, 1, 0))
data$highMED <- ifelse(data$MEDperDay > 100, 1, 0)
data <- na.omit(data[,c(1,5,6,7)])
datagrp <- group_by(data, DIVISION)
points <- summarize(datagrp, Count = n(), Med_MED = median(MEDperDay), Chron_Rate = sum(Chronic)/n(), highMED = 100*sum(highMED)/n())
points <- points[points$Count > 49,]
meanMED <- mean(points$Med_MED)
sdMED <- sd(points$Med_MED)
vert1 <- meanMED + (1 * sdMED)
vert2 <- meanMED + (2 * sdMED)
meanChron <- mean(points$Chron_Rate) 
sdChron <- sd(points$Chron_Rate)
hor1 <- meanChron + (1 * sdChron)
hor2 <- meanChron + (2 * sdChron)

#Plot
p <- ggplot(points, aes(Med_MED, Chron_Rate))
p + geom_vline(aes(xintercept=vert1), color = "blue", linetype = 2) + geom_vline(aes(xintercept=vert2), color = "blue", linetype = 2) + geom_hline(aes(yintercept=hor1), color = "blue", linetype = 2) + geom_hline(aes(yintercept=hor2), color = "blue", linetype = 2) + geom_point(aes(colour = highMED), size = 5) + scale_colour_gradient(high = "red", breaks =c(0,2,4,6), guide = guide_legend(title = "> 100 MEs\nper Day (%)", reverse = T)) + theme_bw() + labs(title = "Variability of Opiate Use by VAMC Division", x = "Median Daily Oral Morphine Equivalents (mg)", y = "Chronic Opiate Use Rate") + annotate("text", x = (vert2+.3), y = .05, colour = "blue", label = "+2 SDs") + annotate("text", x = (vert1+.25), y = .05, colour = "blue", label = "+1 SD") + annotate("text", x = 0, y = (hor1 - .01), colour = "blue", label = "+1 SD") + annotate("text", x = 0, y = (hor2 - .01), colour = "blue", label = "+2 SDs") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  
```